Glucose ingestion before and after resistance training does not augment ribosome biogenesis in healthy moderately trained young adults

### 2023-08-08 manuscript.RMD
updates

\pagebreak
<div custom-style="abstract_style">
## Abstract

@@

<!--*Introduction:*Responses to resistance training (RT) are not uniform across the population, and individuals responding poorly to RT show a blunted ability to produce novel ribosomes [@hammarstrom_benefits_2020; @stec_ribosome_2016]. Recent research solidifies the relationship between increased rates of total RNA and the magnitude of RT-induced adaptations [@hammarstrom_benefits_2020; @hammarstrom_ribosome_2022; @figueiredo_revisiting_2019], with ribosome abundance being associated with UBF protein levels [@hammarstrom_ribosome_2022]. However, 10-15% of individuals undertaking a period of RT showed considerable impaired growth [@alvarez_interindividual_2018; @mann_high_2014; @thalacker-mercer_cluster_2013]. Since modifying RT alone does not convert low-responders to high-responders [@hammarstrom_benefits_2020] other means than RT *per se* is necessary to circumvent this discrepancy. High glucose treatment has been observed to augment rDNA transcription in vitro [@sebata_control_2018], however, its effect on RT-induced adaptations remains quite unexplored.  
*Methods:*Sixteen healthy moderately trained participants were included and randomized to unilateral resistance training with glucose- and placebo supplement. Participants remained fasted overnight receiving glucose/placebo and protein only before and after RT. Resistance training consisted of 3 sets of 10 repetitions maximum unilateral leg press and knee extension. Micro biopsies were sampled pre (T1/T2) and post (T3/T4) the intervention, and maximal unilateral isometric and isokinetic knee extension torque were measured pre (T0), during (days 4, 5, 8, 9) and post (T3/T4, day 13) the intervention.  
*Results:*Despite robust increases in total RNA and rRNA in glucose and placebo (RNA: 26 and 22% respectively, rRNA: 33-43% and 33-41% respectively), there were no differences between supplements (p > 0.05). There were no differences in change of c-Myc, UBF and rpS6 content between treatments (*p* > 0.05). There was indeed a linear relationship between content of UBF and total RNA, where an increase of 1 SD unit in UBF equated to 6.8% increase in total RNA (*p* = 0.002). Lastly, there were no differences between supplements in maximal unilateral isometric knee extension force and isokinetic peak torque during the intervention (p > 0.05) and following the last session (p > 0.05), apart from the 60 d/s test before last RT session (p < 0.05). 
*Conclusion:* Glucose supplement before and after RT does not acutely augment ribosome biogenesis, nor does it enhance muscular performance and recovery after two weeks of heavy RT in moderately trained young individuals. OLD ABSTRACT-->

**Keywords: **

@@

\pagebreak
## Introduction

@@

## Methods

### Materials and methods

@@

### Participants

@@

### Experimental design

@@

### Dietary intervention

@@

### Assessment of muscular strength

@@

### Resistance training protocol

@@

### Sampling of muscle tissue and blood

@@

### Total RNA extraction and quantitative real-time reverse transcription polymerase chain reaction
```{r totrna-extraction}

@@

### Protein extraction and immunoblotting

@@

### Statistics and data analysis

A priori power calculations showed that 20 participants would grant a statistical power of 80% (alpha = 0.05), accounting for an expected dropout of 20%. This power calculation was based on an assumption that the effects of glucose ingestion on ribosome biogenesis (measured as within participant differences in total RNA) may equate to the effects of increasing RT volume from low to moderate [@hammarstrom_benefits_2020] corresponding to an effect size of d =  <!-- Add the effect size here-->.
Total RNA, protein and qPCR data were analyzed by mixed-effects models with fixed effects included as *supplement $\times$ time*. To decrease the risk of Type I errors, random effects were selected from step-wise elimination of terms from the most complex structure (random slopes for time and supplement and their interaction) to less complex. The most complex random effects structure that converged was chosen as the final model[@Matuschek_2017]. Plasma glucose, c-peptide, training volume, and the strength index were analyzed by multiple time-point log-fold change score comparisons, using mixed-effects model with baseline values, time and supplement, and the interaction between time and supplement as fixed effects. These data only supported a random intercept per participant. The linear mixed effects models were fitted with the lmer function from the lme4 package using the lmerTest package to procure p-values (Satterthwaite's method for approximating degrees of freedom)[@bates_fitting_2014; @lmerTest], written for R[@Rcore]. Log-transformed values were expressed as fold-changes in presentations. Descriptive data are presented as mean and (SD). Inferential statistics are presented as means with 95% confidence intervals and *p*-values unless otherwise stated. *p* < 0.05 was considered statistically significant.

## Results

@@

saveRDS(total_rna_stats, "./data/data-gen/rna/total_rna_stats.RDS")

@@

saveRDS(stat_rrna, "./data/data-gen/rna/stat_rrna.RDS")

@@

saveRDS(stat_west, "./data/data-gen/protein/stat_west.RDS")

@@

saveRDS(p_west, "./data/data-gen/protein/p_west.RDS")

@@

corr_incr <- round(100 * (exp(coef(summary(mod))[2,1])-1),1)
saveRDS(corr_incr, "./data/data-gen/protein/corr_incr.RDS")

corrp <- sprintf("%.3f", coef(summary(mod))[2,5])
saveRDS(corrp, "./data/data-gen/protein/corrp.RDS")

@@

corrcoef <- coef(summary(corr.mod))
corrp <- sprintf("%.4f", corrcoef[2,5])

## Percentage increase in RNA per 1 SD uinit increase in UBF using log-model to show relative increse
corr_incr <- round(100 * (exp(coef(summary(corr.lmod))[2,1])-1), 0)

@@

## Discussion

@@

## References

@@

## Figure captions

@@

## Tables

```{r, echo = FALSE}

@@

>>>>

\pagebreak
<div custom-style="abstract_style">
# ABSTRACT

@@

## NEW AND NOTEWORTHY

**Keywords:** Skeletal muscle; hypertrophy; resistance training; glucose; ribosome

@@

\pagebreak
# INTRODUCTION

@@

# MATERIALS AND METHODS

@@

## PARTICIPANTS

@@

## EXPERIMENTAL DESIGN

@@

## DIETARY INTERVENTION

@@

## ASSESSMENT OF MUSCULAR STRENGTH

@@

## RESISTANCE TRAINING PROTOCOL

@@

## SAMPLING OF MUSCLE TISSUE AND BLOOD

@@

## TOTAL RNA EXTRACTION AND REAL-TIME REVERSE TRANSCRIPTION POLYMERASE CHAIN REACTION
```{r totrna-extraction}

@@

## PROTEIN EXTRACTION AND IMMUNOBLOTTING

@@

## STATISTICS AND DATA ANALYSIS

A priori power calculations showed that 20 participants would grant a statistical power of 80% (alpha = 0.05), accounting for an expected dropout of 20%. This power calculation was based on an assumption that the effects of glucose ingestion on ribosome biogenesis (measured as within participant differences in total RNA) may equate to the effects of increasing RT volume from low to moderate [@hammarstrom_benefits_2020] corresponding to an effect size of d =  <!-- Add the effect size here-->.
Total RNA, protein and qPCR data were analyzed by mixed-effects models with fixed effects included as *supplement $\times$ time*. To decrease the risk of Type I errors, random effects were selected from step-wise elimination of terms from the most complex structure (random slopes for time and supplement and their interaction) to less complex. The most complex random effects structure that converged was chosen as the final model[@matuschek_2017_balancing]. Plasma glucose, c-peptide, training volume, and the strength index were analyzed by multiple time-point log-fold change score comparisons, using mixed-effects model with baseline values, time and supplement, and the interaction between time and supplement as fixed effects. These data only supported a random intercept per participant. The linear mixed effects models were fitted with the lmer function from the lme4 package using the lmerTest package to procure p-values (Satterthwaite's method for approximating degrees of freedom)[@bates_fitting_2014; @lmerTest], written for R[@Rcore]. Log-transformed values were expressed as fold-changes in presentations. Descriptive data are presented as mean and (SD). Inferential statistics are presented as means with 95% confidence intervals and *p*-values unless otherwise stated. *p* < 0.05 was considered statistically significant.

# RESULTS

@@

@@

@@

@@

@@

corr_incr <- round(100 * (exp(coef(summary(mod))[2,1])-1),1)

corrp <- sprintf("%.3f", coef(summary(mod))[2,5])

@@

corrcoef <- coef(summary(corr.mod))

corrp <- sprintf("%.3f", coef(summary(mod))[2,5])

corrp <- sprintf("%.4f", corrcoef[2,5])

## Percentage increase in RNA per 1 SD uinit increase in UBF using log-model to show relative increase
corr_incr <- round(100 * (exp(coef(summary(corr.lmod))[2,1])-1), 0)

@@

# DISCUSSION

@@

# APPENDIX

IS THIS REQUIRED? IF NOT, IS IT NECESSARY?

# DATA AVAILABILITY

LINK TO GITHUB

# GRANTS

IS THIS NECESSARY?

# DISCLOSURES

# AUTHOR CONTRIBUTIONS

# References

@@

# FIGURE LEGENDS

@@

# TABLES

```{r baseline table, echo = FALSE}

@@

```{r primer table, echo = FALSE}

### Primer sequence table

## Data
primer.dat <- read_excel("./data/primer.seq2.xlsx")

tab.dat <- primer.dat %>%
  select(Gene, fr, CT.mean, E)

## Flextable
tab.dat %>%
  flextable() %>%
  compose(i = 1, j = c(1, 2, 3, 4), part = "header",
          value = c(as_paragraph("Gene"),
                    as_paragraph("Sequence (forward - reverse)"),
                    as_paragraph("CT mean (SD)"),
                    as_paragraph("E"))) %>%
  #add_header_row(values = c(" ", " ", " ", "Primer sequences"), 
               #  colwidths = c(1, 1, 1, 1)) %>%
  set_caption("Table 2. Primer Sequences") %>%
  width(j = 1, 1.5, unit = "in") %>%
  width(j = 2, 3, unit = "in") %>%
  width(j = 3, 1.5, unit = "in") %>%
  width(j = 4, 0.5, unit = "in")
  #set_table_properties(layout = "autofit")
